---
title: "HW5"
format: html
editor: visual
---

## Setup

```{r}
library(tidyverse)
library(gssr)
library(marginaleffects)
library(broom)

gss2022 <- gss_get_yr(2022)
```

## 5.1

\
This homework has three sections.

Each exercise has two `marginaleffects` outputs: (1) the ATE estimate and (2) the ATT/ATU estimates.

You will have to reproduce these estimates *without* using `marginaleffects`. There are a couple of ways to do this, but you will probably end up using the `predict()` function (from base R), or the `augment()` function (from the `broom` package).

Both functions have an argument called `newdata`, which you can use doing something similar to this toy example:

```{r}
ols <- lm(mpg ~ disp + am, data = mtcars)

new_am0 <- mtcars |> 
  mutate(am = 0)

new_am1 <- mtcars |> 
  mutate(am = 1)

p0 <- predict(ols, newdata = new_am0) ## predictions for am == 0
p1 <- predict(ols, newdata = new_am1) ## predictions for am == 1
```

## 5.2 Linear Regression

data

```{r}
d <- gss2022 |> 
  select(tvhours, degree, madeg, padeg) |> 
  mutate(pardeg = pmax(madeg, padeg, na.rm = TRUE),
         college = if_else(degree >= 3, 1L, 0L),
         parcol = if_else(pardeg >= 3, 1L, 0L)) |>
  select(tvhours, college, parcol) |> 
  drop_na()
```

## 5.2.1

Additive link function, no interactions

```{r}
mod1 <- lm(tvhours ~ college + parcol, data = d)

# ATE estimate
avg_slopes(mod1, variables = "college") |> 
  tidy()
```

ANSWER:

```{r}
d521_1 = mutate(d, college = 1)
d521_0 = mutate(d, college = 0)

p521_1 = predict(mod1, newdata = d521_1)
p521_0 = predict(mod1, newdata = d521_0)

mean(p521_1 - p521_0)
```

```{r}
# ATT/ATU estimate
avg_slopes(
  model = mod1, 
  variables = "college",
  by = "college" # separately by treatment group
) |> 
  tidy()
```

ANSWER:

```{r}
mean(predict(mod1, newdata = d521_1)[d$college == 0]) - mean(predict(mod1, newdata = d521_0 )[d$college == 0])

mean(predict(mod1, newdata = d521_1)[d$college == 1]) - mean(predict(mod1, newdata = d521_0 )[d$college == 1])
```

## 5.2.2

Additive link function, with interactions

```{r}
mod2 <- lm(tvhours ~ college * parcol, data = d)

# ATE estimate
avg_slopes(mod2, variables = "college") |> 
  tidy()
```

ANSWER:

```{r}
p522_1 = predict(mod2, newdata = d521_1)
p522_0 = predict(mod2, newdata = d521_0)

mean(p522_1 - p522_0)
```

```{r}
# ATT/ATU estimate
avg_slopes(
  model = mod2, 
  variables = "college",
  by = "college" # separately by treatment group
) |> tidy()
```

ANSWER:

```{r}
mean(predict(mod2, newdata = d521_1)[d$college == 0]) - mean(predict(mod2, newdata = d521_0 )[d$college == 0])

mean(predict(mod2, newdata = d521_1)[d$college == 1]) - mean(predict(mod2, newdata = d521_0 )[d$college == 1])
```

## 5.3 Poisson Regression

data

```{r}
d <- gss2022 |>
  filter(wrkstat == 1) |> # full time workers
  select(realrinc, degree, madeg, padeg, sex, age) |> 
  mutate(pardeg = pmax(madeg, padeg, na.rm = TRUE),
         college = if_else(degree >= 3, 1L, 0L),
         parcol = if_else(pardeg >= 3, 1L, 0L),
         female = if_else(sex == 2, 1L, 0L),
         realrinc = floor(realrinc)) |>             # integer
  select(realrinc, college, parcol, female, age) |> 
  drop_na()
```

## 5.3.1

Using the log-counts, no interactions

```{r}
qp1 <- glm(realrinc ~ college + (parcol + female + age + I(age^2)), 
           data = d,
           family = "quasipoisson")

avg_slopes(qp1,
           variables = "college",
           type = "link") |> 
  tidy()
```

ANSWER:

```{r}
d531_1 = mutate(d, college = 1)
d531_0 = mutate(d, college = 0)

p531_1 = predict(qp1, newdata = d531_1)
p531_0 = predict(qp1, newdata = d531_0)

mean(p531_1 - p531_0)
```

```{r}
avg_slopes(qp1,
           variables = "college",
           type = "link",
           by = "college") |> # separately by treatment group
  tidy()
```

ANSWER:

```{r}
mean(predict(qp1, newdata = d531_1)[d$college == 0]) - mean(predict(qp1, newdata = d531_0 )[d$college == 0])

mean(predict(qp1, newdata = d531_1)[d$college == 1]) - mean(predict(qp1, newdata = d531_0 )[d$college == 1])
```

## 5.3.2

Non-linear response, no interactions

```{r}
avg_slopes(qp1,
           variables = "college",
           type = "response") |> 
  tidy()
```

ANSWER:

```{r}
d531_1 = mutate(d, college = 1)
d531_0 = mutate(d, college = 0)

p531_1 = predict(qp1, newdata = d531_1, type = "response")
p531_0 = predict(qp1, newdata = d531_0, type = "response")

mean(p531_1 - p531_0)
```

```{r}
avg_slopes(qp1,
           variables = "college",
           type = "response",
           by = "college") |> # separately by treatment group
  tidy()
```

ANSWER:

```{r}
mean(predict(qp1, newdata = d531_1, type = "response")[d$college == 0]) - mean(predict(qp1, newdata = d531_0 , type = "response")[d$college == 0])

mean(predict(qp1, newdata = d531_1, type = "response")[d$college == 1]) - mean(predict(qp1, newdata = d531_0, type = "response" )[d$college == 1])
```

## 5.3.3

Using the log-counts, with interactions

```{r}
qp2 <- glm(realrinc ~ college * (parcol + female + age + I(age^2)), 
           data = d,
           family = "quasipoisson")

avg_slopes(qp2,
           variables = "college",
           type = "link") |> 
  tidy()
```

ANSWER:

```{r}
p533_1 = predict(qp2, newdata = d531_1)
p533_0 = predict(qp2, newdata = d531_0)

mean(p533_1 - p533_0)
```

```{r}
avg_slopes(qp2,
           variables = "college",
           type = "link",
           by = "college") |> # separately by treatment group
  tidy()
```

ANSWER:

```{r}
mean(predict(qp2, newdata = d531_1)[d$college == 0]) - mean(predict(qp2, newdata = d531_0 )[d$college == 0])

mean(predict(qp2, newdata = d531_1)[d$college == 1]) - mean(predict(qp2, newdata = d531_0 )[d$college == 1])
```

## 5.3.4

Non-linear response, with interactions

```{r}
avg_slopes(qp2,
           variables = "college",
           type = "response") |> 
  tidy()
```

ANSWER:

```{r}
p534_1 = predict(qp2, newdata = d531_1, type = "response")
p534_0 = predict(qp2, newdata = d531_0, type = "response")

mean(p534_1 - p534_0)
```

```{r}
avg_slopes(qp2,
           variables = "college",
           type = "response",
           by = "college") |> # separately by treatment group
  tidy()
```

ANSWER:

```{r}
mean(predict(qp2, newdata = d531_1, type = "response")[d$college == 0]) - mean(predict(qp2, newdata = d531_0, type = "response" )[d$college == 0])

mean(predict(qp2, newdata = d531_1 , type = "response")[d$college == 1]) - mean(predict(qp2, newdata = d531_0, type = "response" )[d$college == 1])
```

## 5.4 Logistic Regression

data

```{r}
d <- gss2022 |>
  select(abany, degree, madeg, padeg, sex, age) |> 
  mutate(pardeg = pmax(madeg, padeg, na.rm = TRUE),
         college = if_else(degree >= 3, 1L, 0L),
         parcol = if_else(pardeg >= 3, 1L, 0L),
         female = if_else(sex == 2, 1L, 0L),
         abany = if_else(abany == 1, 1L, 0L)) |>
  select(abany, college, parcol, female, age) |> 
  drop_na()
```

## 5.4.1

Using log-odds, no interactions

```{r}
lr1 <- glm(abany ~ college + (parcol + female + age + I(age^2)),
          data = d,
          family = binomial)

# ATE estimate
avg_slopes(lr1,
           variables = "college",
           type = "link") |> 
  tidy()
```

ANSWER:

```{r}
d541_1 = mutate(d, college = 1)
d541_0 = mutate(d, college = 0)

p541_1 = predict(lr1, newdata = d541_1)
p541_0 = predict(lr1, newdata = d541_0)

mean(p541_1 - p541_0)
```

```{r}
avg_slopes(lr1,
           variables = "college",
           by = "college",
           type = "link") |> 
  tidy()
```

ANSWER:

```{r}
mean(predict(lr1, newdata = d541_1)[d$college == 0]) - mean(predict(lr1, newdata = d541_0 )[d$college == 0])

mean(predict(lr1, newdata = d541_1)[d$college == 1]) - mean(predict(lr1, newdata = d541_0 )[d$college == 1])
```

## 5.4.2

Using non-linear response (aka probabilities), no interactions

```{r}
# ATE estimate
avg_slopes(lr1,
           variables = "college",
           type = "response") |> 
  tidy()
```

ANSWER:

```{r}
p542_1 = predict(lr1, newdata = d541_1, type = "response")
p542_0 = predict(lr1, newdata = d541_0, type = "response")

mean(p542_1 - p542_0)
```

```{r}
avg_slopes(lr1,
           variables = "college",
           by = "college",
           type = "response") |> 
  tidy()
```

ANSWER:

```{r}
mean(predict(lr1, newdata = d541_1, type = "response")[d$college == 0]) - mean(predict(lr1, newdata = d541_0, type = "response" )[d$college == 0])

mean(predict(lr1, newdata = d541_1, type = "response")[d$college == 1]) - mean(predict(lr1, newdata = d541_0, type = "response" )[d$college == 1])
```

## 5.4.3

Using log-odds, with interactions

```{r}
lr2 <- glm(abany ~ college * (parcol + female + age + I(age^2)),
          data = d,
          family = binomial)

# ATE estimate
avg_slopes(lr2,
           variables = "college",
           type = "link") |> 
  tidy()
```

ANSWER:

```{r}
p543_1 = predict(lr2, newdata = d541_1)
p543_0 = predict(lr2, newdata = d541_0)

mean(p543_1 - p543_0)
```

```{r}
avg_slopes(lr2,
           variables = "college",
           by = "college",
           type = "link") |> 
  tidy()
```

ANSWER:

```{r}
mean(predict(lr2, newdata = d541_1)[d$college == 0]) - mean(predict(lr2, newdata = d541_0 )[d$college == 0])

mean(predict(lr2, newdata = d541_1)[d$college == 1]) - mean(predict(lr2, newdata = d541_0 )[d$college == 1])
```

## 5.4.4

Using non-linear response (aka probabilities), with interactions

```{r}
# ATE estimate
avg_slopes(lr2,
           variables = "college",
           type = "response") |> 
  tidy()
```

ANSWER:

```{r}
p544_1 = predict(lr2, newdata = d541_1, type = "response")
p544_0 = predict(lr2, newdata = d541_0, type = "response")

mean(p544_1 - p544_0)
```

```{r}
avg_slopes(lr2,
           variables = "college",
           by = "college",
           type = "response") |> 
  tidy()
```

ANSWER:

```{r}
mean(predict(lr2, newdata = d541_1, type = "response")[d$college == 0]) - mean(predict(lr2, newdata = d541_0, type = "response" )[d$college == 0])

mean(predict(lr2, newdata = d541_1, type = "response")[d$college == 1]) - mean(predict(lr2, newdata = d541_0, type = "response" )[d$college == 1])
```
